---
title: "Week4_Assignment"
author: "Chen Liu"
date: "2/8/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assignment helps understanding stationarity and seasonality of linear models for time series

Exercise 7 on page 126 of the Textbook

```{r,warning=FALSE,include=FALSE}
library(tseries)
library(forecast)
library(TSA)
```

Load in the data
```{r}
datapath<-"/Users/me/Desktop/MSCA/FinancialAnalytics/week4"
da<-read.table(file=paste(datapath,"q-jnj-earns-9211.txt",sep="/"),header=T)
head(da)
dim(da)
plot(da$earns)
```
almost linear growth in earns, so take log transformation
```{r}
earns<-log(da$earns)
plot(earns,type = 'l',main='Log transformation of earnings')
```

Still not stationary, take the difference of the log earnings.
```{r}
diff.earns<-diff(earns)
plot(diff.earns,type = 'l')
acf(diff.earns)
```

It now looks more stationary, and shows seasonality.
```{r}
#consider seasonality
s.diff<-diff(earns,4)
#consider both seasonality and regular difference
d.s<-diff(s.diff)
par(mfrow=c(2,2))
acf(earns)
acf(diff.earns)
acf(s.diff)
acf(d.s)

```

The ACF of the log of earnings shows long memory, which is not stationary. Taking the difference of the log earnigs shows seasonality.
Taking regular and seasonal difference shows exponential decay.

Make the time series plots after differencing.
```{r}
# Obtain time plots
par(mfcol=c(3,1))
plot(diff.earns,xlab='year',ylab='diff',type='l')
points(diff.earns,pch='c1',cex=0.7)
plot(s.diff,xlab='year',ylab='sea-diff',type='l')
points(s.diff,pch='c2',cex=0.7)
plot(d.s,xlab='year',ylab='dd',type='l')
points(d.s,pch='c1',cex=0.7) 

# adf test
adf.test(diff.earns)
adf.test(s.diff)
adf.test(d.s)
```
Regular differencing (top) removed growth and stressed seasonality. The ACF of that is high for lags which are multiles of 4. ACF decays slowly. Looks like seasonality at 4.
Only seasonal differencing (middle) removed seasonality, but did not get stationarity.

Both differences (bottom) removed both seasonality and nonstationarity.

Estimate the model ARIMA(0,1,1) 
```{r}
m1=arima(earns,order=c(0,1,1),seasonal=list(order=c(0,1,1),period=4))
m1
```

```{r}
tsdiag(m1,gof=20) 
```
```{r}
Box.test(m1$residuals,lag=12,type='Ljung')
```
```{r}
pp1=1-pchisq(6.204,10)
pp1
```
The test of residuals shows the model fits well. 

My model is : (1-0.3223B)(1-0.2175B4)ϵt,σϵ2= 0.0011

Refit the model using data from 1992 to 2008
```{r}
da2<-earns[1:68]
m1<-arima(da2,order=c(0,1,1),seasonal=list(order=c(0,1,1),period=4))
m1

```
My refitted model is : (1-0.3419B)(1-0.1849B4)ϵt,σϵ2= 0.001002

Perform 1 to 10 step forecasts of earnings and obtain a forecast plot.
```{r}
par(mfrow=c(1,1))
pm1<-predict(m1,10)
pred<-pm1$pred
se<-pm1$se
ko<-da$earns # actual observations
fore<-exp(pred+se^2/2) #point forecasts, delogged
v1<-exp(2*pred+se^2)*(exp(se^2)-1)
s1<-sqrt(v1) # std of the forecast error
eps<-ko[60:78]
length(eps)
tdx<-(c(1:19)+3)/4+2006
upp<-c(ko[68],fore+2*s1) # upper band (+2*std)
low<-c(ko[68],fore-2*s1) # lower band (-2*std)
plot(tdx,eps,xlab='year',ylab='earnings',type='l',ylim=c(min(low,eps)-0.3,max(upp,eps)+0.3))
points(tdx[10:19],fore,pch='*')
lines(tdx[9:19],upp,lty=2)
lines(tdx[9:19],low,lty=2)
points(tdx[10:19],ko[69:78],pch='o',cex=0.7)
```
Dashed lines - confidence bounds.
Circles - actual earnings.
Asterisks - predicted values.







