---
title: "Assignment3_ChenLiu"
author: "Chen Liu"
date: "1/26/2020"
output: pdf_document
---

```{r setup,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tseries)
library(forecast)
library(TSA)
```

1. Exercise 2 on page 125
Use the file m-dec125910-6111.txt from the website of the Author


```{r}
datapath<-"/Users/me/Desktop/MSCA/FinancialAnalytics/week3"
da<-read.table(paste(datapath,"m-dec125910-6111.txt",sep="/"),header=T)
head(da)
dim(da)
```

1.1 For CRSP portfolios of Decile 2, 10 returns test null hypothesis that the first 12 lags of autocorrelations equal zero with 5% level

```{r,Decile2,warning=FALSE}
dc2<-da$dec2
dc2.ts<-ts(dc2,frequency=12,start=c(1961,1))
lndc2<-log(dc2+1)
lndc2.ts<-ts(lndc2,frequency=12)
par(mfcol=c(1,1))
plot(dc2.ts,xlab='year',ylab='returns')
title(main='Simple returns')
acf(dc2,lag=24)
#plot log returns
plot(lndc2.ts,xlab='year',ylab='returns')
title(main='log returns')
acf(lndc2,lag=24)
```

Test the null hypothesis that the first 12 lags of autocorrelations of decile2 equal zero with 5% level.
```{r}
Box.test(dc2,lag=12,type='Ljung')
Box.test(lndc2,lag=12,type='Ljung')
```
We see that Box-Ljung test cannot reject H0 of no correlations first 12 lags of decile2 returns with 5% level.

```{r,Decile 10}
dc10<-da$dec10
dc10.ts<-ts(dc10,frequency=12,start=c(1961,1))
lndc10<-log(dc10+1)
lndc10.ts<-ts(lndc10,frequency=12)
par(mfcol=c(1,1))
plot(dc10.ts,xlab='year',ylab='returns')
title(main='Simple returns')
acf(dc10,lag=24)
#plot log returns
plot(lndc10.ts,xlab='year',ylab='returns')
title(main='log returns')
acf(lndc10,lag=24)
Box.test(dc10,lag=12,type='Ljung')
Box.test(lndc10,lag=12,type='Ljung')
```
We see that Box-Ljung test results return both p-values that are significant at the 5% level. Therefore, we can reject the null hypothesis and say that there exists serial correlations in the monthly simple returns of the Decile 10 data.

1.2 Fit ARMA model for returns of Decile 2, perform model checking and write down the fitted model
```{r}
acf(dc2)
pacf(dc2)
eacf(dc2)
auto.arima(dc2)
arma1<-arima(dc2,order=c(0,0,1))
arma11<-arima(dc2,order=c(1,0,1))
arma11
arma1
```

```{r}
Box.test(arma1$residuals,lag=12,type='Ljung')
Box.test(arma11$residuals,lag=12,type='Ljung')
1-pchisq(9.4993,11)
1-pchisq(7.8567,10)
tsdiag(arma11)
tsdiag(arma1)
```
I choose ARMA(0,1) model, which has lower AIC. And the test shows that the residual is not autocorrelated.


1.3 Use the fitted model to produce 1- to 12-step ahead forecasts of the series and the associated standard errors of forecasts.
```{r}
prd <- predict(arma1,12)
cbind(Predicted=as.vector(prd$pred),Standard.Error=as.vector(prd$se))
```


2.Consider logs of monthly yields of Moody’s AAA bonds. Fit a time series model, do the model checking.
```{r}
Moody<-read.table(paste(datapath,"m-aaa-1911.txt",sep="/"),header=T)
dim(Moody)
lny<-diff(log(Moody$yield))
plot(lny,xlab='year',ylab='yield')
title(main='Monthly yield')
#Test stationary
adf.test(lny)
```
The data set passed the stationary test.
```{r}
acf(lny, lag =30)
pacf(lny, lag =30)
```

The time series of monthly yield is not stationary.
```{r}
eacf(lny)
auto.arima(lny)
```
Based on the eacf and auto.arima, I have several candidates:ARIMA(2,0,3), ARIMA(0,0,3), ARIMA(3,0,2)
```{r}
model1<-arima(lny,order=c(2,0,3))
model2<-arima(lny,order=c(0,0,3))
model3<-arima(lny,order=c(3,0,2))

AIC1<-model1$aic
AIC2<-model2$aic
AIC3<-model3$aic

df <- data.frame(models = c("ARMA(2,3)","MA(3)","ARMA(3,2)"),AIC=c(AIC1,AIC2,AIC3))
knitr::kable(df, caption = "AICs of the Three Candidate Models")
```
Based on the AIC value, we choose model MA(3).

Model checking
```{r}
Box.test(model2$residuals,lag=12,type='Ljung')
1-pchisq(6.9494,9)
tsdiag(model2)
```

Because the p-value is larger than 5%, we cannot reject the null hopytheis, the residual is not autocorrelated. Therefore, our model works.





















